Spin canonical divisors of log stacky curves

### 2015-06-25 stacky_write_up/spin_cring.tex
added g > 1 induction

\ssec{Inductive Lemmata: Adding Points}
\label{ssec:add_pts}
In this section, we prove various inductive statements under the following setup:
Let $\halfcan'$ be $\mathbb{Q}$-divisor of $X$ (and often a log half canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that minimally generate\todo{check minimal generation actually works} $R_\halfcan$ over $\halfcan$' and initial terms of relations that minimally generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Grobner Basis).  The proofs of these Lemmas all follow similar patterns, so we will present a few \todo{give the precise number we show} and leave the rest to the reader.

@@

\begin{lem}
\label{lem:deg1-sat-ind}
Let $X$ be a genus $g$ curve with a $\mathbb{Q}$-divisor $\halfcan'$
satisfying $h^0(X, \lfloor{\halfcan'}\rfloor)\ge 1$. Suppose $P$ is not a base-point of $d\halfcan'$ for all $d\in \mathbb{N}^+$ meaning that $R_{\halfcan'}$ has generically chosen generators (in the necessary degrees)
$u, x_1, \ldots, x_m$ with $\ord_P(x_i)=0\; \forall x_i$, $\ord_P(u) = 0$, \todo{Aaron: Instead of having $\deg u = 0$, don't we only need that $\ord_P(u) = \lfloor D|_P \rfloor$?} and $\deg(u)= 1$ when viewed as sections of $\so_D$.  Suppose $u, x_1, \ldots, x_m$ generate $R_D$ in degree at most $\bar{d}$ with relations generated
in degree at most $\bar{\tau}$.  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbb{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \text{ for all } d \in \mathbb{
	N} \text{ such that } d \ge \frac{\beta}{\alpha}.
\end{equation}
\label{eqn:deg1-sat-ind}
Let 
\[
	0<\frac{c_1}{d_1}<\ldots<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the positive best lower approximations of $\frac{\alpha}{\beta}$.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1, \ldots, y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$

\item If $I$ and $I'$ are the ideal of relations of $R_D$ and $R_{D'}$ respectively, then

@@

In particular, if $\frac{\alpha}{\beta}\le 1$, then $I$ is minimally generated by elements with leading terms of the form $y_i x_j$ and $y_i y_j$.
%\item If $\gin_\prec(I)=\initial_\prec(I)$ then $\gin_\prec(I')=\initial_\prec(I')$.
\item A Grobner basis with the initial terms $y_i x_j$, $y_i y_j$,
	and possibly $(y_{\lfloor \frac{\alpha}{\beta} \rfloor})^{y_{\lfloor
	\frac{\alpha}{\beta} \rfloor} + 1}$ as shown above minimally
generates $I'$.
\end{itemize}
\end{lem}

@@

\begin{lem}
\label{lem:sat-two-induction-high-genus}
Let $(\sx, \Delta, \halfcan_\sx)$ and $(\sx', \Delta, \halfcan_{\sx'})$ be log stacky curves with coarse space $(X, \Delta, L_X)$ and signatures $(g; e_1, \ldots, e_r; \delta)$ and $(g, e_1, \ldots, e_{r- 1}, \delta)$ respectively, with $g > 0,$ and, if $g = 1$ then $\deg 3L \geq 2$.

@@

\begin{defn}
\label{defn:admissible}
Let $(\sx', 0, L'),(\sx, 0, L)$ be tame, 
separably rooted log stacky spin
curves, both ramified over ramified at $Q_1, \ldots, Q_r$. Suppose
there is a birational map $\sx \rightarrow \sx'$, and let $J \subset
\{1, \ldots, r\}$. Suppose $e_i'+ 2 \chi_J (i) = e_i$ where

@@

\begin{proof}
Let $(\sx, 0, L)$ and $(\sx', 0, L')$ be the two log stacky spin curves in 
question, with coarse spaces $X = X'$. 
\todo{possibly generalize these to arbitrary log stacky $\Delta \neq 0$ spin curves}
Define

@@

\begin{cor}
\label{cor:raise-stacky-order}
Suppose $(\sx',L',J')$ is admissible with signature $\sigma' = (e_1',\ldots,e_r')$ with $J' = \{t,t+1,\ldots,r\}$ and $e_1' \leq e_2' \leq \cdots \leq e_t' = e_{t+1}' =\cdots = e_r'$
so that $(\sx',\Delta, L)$ satisfies the conditions of Theorem ~\ref{thm:ram_order_ind} and Theorem ~\ref{thm:general_generators_relations}. Then, for any spin curve $(\sx,\Delta,L)$ so that $\sx$ and $\sx'$ have the same coarse space $X = X'$ with the same set of ramified points, and $\sx$ has signature $(g;e_1,\ldots, e_r;\delta)$ with $e_1 \leq e_2 \leq \cdots e_r$ so that
$$
e_i \begin{cases}
	= e_i' &\text{ if }i \notin J\\
	\leq e_i' &\text{ if } i \in J
\end{cases}$$
then Theorem ~\ref{thm:general_generators_relations} holds for $(\sx,\Delta,L)$
\end{cor}
\begin{proof}
For $t \leq i \leq r$. Let $(\sx_i,\Delta, L_i)$ be the log stacky spin curve with the same coarse space as $L'$ and with the same stacky points as $(\sx',\Delta,L),$ having signature $(g;e_1, \ldots e_{i-1}, e_i, e_i, \ldots, e_i;\delta),$ and let 
$J_i = \{i,\ldots, r\}.$ Note that $(\sx_k,L_k,J_0) = (\sx',L',J')$ and $(\sx_r ,L_r ,J_r) = (\sx, L, J)$.

@@

\begin{proof}
By Lemma ~\ref{lem:generation_5} and Proposition ~\ref{prop:relation_11}, we are done in the case that there are no stacky points. 
\end{proof}

@@

We next check $(\sx',\{3,3\})$ is admissible. We clearly have two generators $x_3, y_3$ in degree 3 with a pole of order $\frac{3- 1}{2}$ by construction. Hence, (Ad-i) holds. We next check (Ad-ii) for the point $P_1$, as the case of $P_2$ is symmetric. Here, by construction
$$
\frac{-\ord_{P_1}(z)}{\ord(z)} = \begin{cases}
	0 &\text{ if }z \in \{u, y_3\}\\
	\frac{1}{4} &\text{ if }z = y_4
\end{cases}.$$
Since $0 < \frac{1}{4} < \frac{1}{3}$, (Ad-ii) holds.
Finally, to check (Ad-iii) note that $\max_{d \geq 0}S_{\sigma,J}(i,d) = 0.$ Therefore, $\deg \lfloor 5L \rfloor  = 2 > 1 = 2 \cdot 1 - 1 + 0$.
\end{example}

>>>>

\ssec{Inductive Lemmata: Adding Points}
\label{ssec:add_pts}
In this section, we prove various inductive statements under the following setup:
Let $\halfcan'$ be $\mathbb{Q}$-divisor of $X$ (and often a log half-canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that minimally generate\todo{check minimal generation actually works} $R_\halfcan$ over $\halfcan$' and initial terms of relations that minimally generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Grobner Basis).  The proofs of these Lemmas all follow similar patterns, so we will present a few \todo{give the precise number we show} and leave the rest to the reader.

@@

\begin{lem}
\label{lem:deg1-sat-ind}
Let $X$ be a genus $g$ curve with a $\mathbb{Q}$-divisor $\halfcan'$
satisfying $h^0(X, \lfloor{\halfcan'}\rfloor)\ge 1$. Suppose $P$ is not a base-point of $d\halfcan'$ for all $d\in \mathbb{N}^+$ meaning that $R_{\halfcan'}$ has generically chosen generators (in the necessary degrees)
$u, x_1, \ldots, x_m$ with $\ord_P(x_i)=0\; \forall x_i$, $\ord_P(u) = 0$, \todo{Aaron: Instead of having $\deg u = 0$, don't we only need that $\ord_P(u) = \lfloor D|_P \rfloor$?} and $\deg(u)= 1$ when viewed as sections of $\so_D$.  Suppose $u, x_1, \ldots, x_m$ generate $R_D$ in degree at most $\bar{d}$ with relations generated
in degree at most $\bar{\tau}$.  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbb{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \text{ for all } d \in \mathbb{
	N} \text{ such that } d \ge \frac{\beta}{\alpha}.
\end{equation}
\label{eqn:deg1-sat-ind}
Let 
\[
	0<\frac{c_1}{d_1}<\ldots<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the positive best lower approximations of $\frac{\alpha}{\beta}$.
Then 
\begin{enumerate}
\item[(a)] $R_{D'}$ is generated over $R_D$ by elements $y_1, \ldots, y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$

\item[(b)] If $I$ and $I'$ are the ideal of relations of $R_D$ and $R_{D'}$ respectively, then

@@

In particular, if $\frac{\alpha}{\beta}\le 1$, then $I$ is minimally generated by elements with leading terms of the form $y_i x_j$ and $y_i y_j$.
%\item If $\gin_\prec(I)=\initial_\prec(I)$ then $\gin_\prec(I')=\initial_\prec(I')$.
\item[(c)] A Grobner basis with the initial terms $y_i x_j$, $y_i y_j$,
	and possibly $(y_{\lfloor \frac{\alpha}{\beta} \rfloor})^{y_{\lfloor
	\frac{\alpha}{\beta} \rfloor} + 1}$ as shown above minimally
generates $I'$.
\end{enumerate}
\end{lem}

@@

\begin{lem}
\label{lem:sat-two-induction-high-genus}
Let $(\sx, \Delta, \halfcan_\sx)$ and $(\sx', \Delta, \halfcan_{\sx'})$ be log spin curves with coarse space $(X, \Delta, L_X)$ and signatures $(g; e_1, \ldots, e_r; \delta)$ and $(g, e_1, \ldots, e_{r- 1}, \delta)$ respectively, with $g > 0,$ and, if $g = 1$ then $\deg 3L \geq 2$.

@@

\begin{defn}
\label{defn:admissible}
Let $(\sx', 0, L'),(\sx, 0, L)$ be log spin
curves, both ramified over ramified at $Q_1, \ldots, Q_r$. Suppose
there is a birational map $\sx \rightarrow \sx'$, and let $J \subset
\{1, \ldots, r\}$. Suppose $e_i'+ 2 \chi_J (i) = e_i$ where

@@

\begin{proof}
Let $(\sx, 0, L)$ and $(\sx', 0, L')$ be the two log spin curves in 
question, with coarse spaces $X = X'$. 
\todo{possibly generalize these to arbitrary log  $\Delta \neq 0$ spin curves}
Define

@@

\begin{cor}
\label{cor:raise-stacky-order}
Suppose $(\sx',L',J')$ is admissible with signature $\sigma' = (e_1',\ldots,e_r')$ with $J' = \{t,t+1,\ldots,r\}$ and $e_1' \leq e_2' \leq \cdots \leq e_t' = e_{t+1}' =\cdots = e_r'$
so that $(\sx',\Delta, L)$ satisfies the conditions of Theorem ~\ref{thm:ram_order_ind} and Theorem ~\ref{thm:general_generators_relations}. Then, for any spin curve $(\sx,\Delta,L)$ so that $\sx$ and $\sx'$ have the same coarse space $X = X'$ with the same set of ramified points, and $\sx$ has signature $(g;e_1,\ldots, e_r;\delta)$ with $e_1 \leq e_2 \leq \cdots e_r$ so that
$$
e_i \begin{cases}
	= e_i' &\text{ if }i \notin J\\
	\leq e_i' &\text{ if } i \in J
\end{cases}$$
then Theorem ~\ref{thm:general_generators_relations} holds for $(\sx,\Delta,L)$
\end{cor}
\begin{proof}
For $t \leq i \leq r$. Let $(\sx_i,\Delta, L_i)$ be the log spin curve with the same coarse space as $L'$ and with the same stacky points as $(\sx',\Delta,L),$ having signature $(g;e_1, \ldots e_{i-1}, e_i, e_i, \ldots, e_i;\delta),$ and let 
$J_i = \{i,\ldots, r\}.$ Note that $(\sx_k,L_k,J_0) = (\sx',L',J')$ and $(\sx_r ,L_r ,J_r) = (\sx, L, J)$.

@@

\begin{proof}
Let $\halfcan_X \in \di X$ satisfy $2\halfcan_X + 2 K_X + \Delta$. By Lemma ~\ref{lem:generation_5} and Proposition ~\ref{prop:relation_11}, the theorem holds for $(\sx, \Delta, \halfcan_X)$. Next, consider a divisor of the form $\halfcan' = \halfcan_X + \frac{1}{3}P_i$. Then, the theorem holds for $(\sx,\Delta, \halfcan')$ by Lemma ~\ref{lem:sat-two-induction-high-genus} if $P_i$ is a basepoint of $\halfcan'$ or by Lemma ~\ref{deg1-sat-ind} if $P_i$ is not a basepoint of $L'$. Note, in the case $P$ is a not basepoint of $\halfcan'$, Equation ~\ref{eqn:deg1-sat-ind} holds by Riemann Roch, and the fact that $d \geq 3$.

In this case, we claim $(\sx',L',\{1\})$ is admissible. Indeed, (Ad-i) holds by part (a) of Lemmas ~\ref{lem:sat-two-induction-high-genus} and ~\ref{deg1-sat-ind}, (Ad-ii) holds by part (b) of Lemmas ~\ref{lem:sat-two-induction-high-genus} and ~\ref{deg1-sat-ind}, and (Ad-iii) holds as $5\deg L \geq 5(g-1)> 2g-1.$ Therefore, by Lemma ~\ref{lem:n-3s-admissible}, Any triple of the form $(\sx',L',\{1,\ldots, r\})$ is admissible, with $e_1 = \cdots = e_r = 3$. Hence, by Corollary ~\ref{cor:raise-stacky-order}, this theorem holds for all log spin curves $(\sx, L, J).$ 
\end{proof}

@@

We next check $(\sx',\{1,2\})$ is admissible. We clearly have two generators $x_3, y_3$ in degree 3 with a pole of order $\frac{3- 1}{2}$ by construction. Hence, (Ad-i) holds. We next check (Ad-ii) for the point $P_1$, as the case of $P_2$ is symmetric. Here, by construction
$$
\frac{-\ord_{P_1}(z)}{\ord(z)} = \begin{cases}
	0 &\text{ if }z \in \{u, y_3\}\\
	\frac{1}{4} &\text{ if }z = y_4
\end{cases}.$$
Since $0 < \frac{1}{4} < \frac{1}{3}$, (Ad-ii) holds.
Finally, to check (Ad-iii) note that $\max_{d \geq 0}S_{\sigma,J}(i,d) = 0.$ Therefore, $\deg \lfloor 5L \rfloor  = 2 > 1 = 2 \cdot 1 - 1 + 0$.
\end{example}
